### 2022-04-27 inst/analysis/1-graphs.Rmd
add code to generate boxplots

---
title: "Graphs"
author: "Beatriz Milz"
date: "3 de novembro de 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dpi = 600)
```

# Load packages
```{r}
library(ggplot2)
devtools::load_all()
```

## Prepare dataset
```{r}
sampling_data  <- sampling_data_long %>%
  dplyr::mutate(sampling_point = forcats::as_factor(sampling_sites)) %>%
  dplyr::left_join(variable_names) %>%
  dplyr::mutate(
    variable_name_abrev  = dplyr::case_when(
      variable == "water_temperature" ~ "Water Temperature (WT)",
      variable == "dissolved_oxygen" ~ "Dissolved Oxygen (DO)",
      variable == "ph" ~ "pH",
      variable == "electrical_conductivity" ~ " Electrical Conductivity (EC)",

      TRUE ~ paste0(variable_name, " (", variable, ")")
    )
  )

dplyr::glimpse(sampling_data)
```

## WT/DO/pH/EC

### Water temperature
```{r}
boxplot_wt <- create_boxplot(variavel_y = "water_temperature",
                             legenda_y =  "Temperature\n (°C)")

boxplot_wt
```

### Dissolved oxygen
```{r}
boxplot_od <- create_boxplot(
               variavel_y = "dissolved_oxygen",
               legenda_y =  "Dissolved oxygen\n (mg/L)")

boxplot_od
```

### pH
```{r}
boxplot_ph <- create_boxplot(
  variavel_y = "ph",
  legenda_y =  "pH",
  esconder_legenda_x = TRUE,
  esconder_legenda_fill = FALSE
)

boxplot_ph
```

### Electrical conductivity

```{r}
boxplot_ce <- create_boxplot(
               variavel_y = "electrical_conductivity",
               legenda_y =  "Electrical conductivity\n (μS/cm)",
               esconder_legenda_x = TRUE)

boxplot_ce
```

## Nutrients

```{r}
boxplot_nutri <- sampling_data %>%
  dplyr::filter(variable %in% c("TN", "NO2", "NO3", "NH4", "TP", "TDP", "SRP", "CHLA")) %>%
  dplyr::mutate(
    variable = forcats::fct_relevel(variable ,
                                    "TN", "NO2", "NO3", "NH4", "TP", "TDP", "SRP", "CHLA")
  ) %>%
  ggplot() +
  geom_boxplot(aes(x = season, y = result_variable, fill = sampling_point)) +
  scale_fill_brewer(palette = "Greys", direction = -1) +
  labs(x = "Season",
       y = "Concentration (μg/L)",
       fill = "Sampling point") +
  theme_bw() +
  facet_wrap( ~ variable_name_abrev, scales = "free_y", ncol = 2) +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))

boxplot_nutri

ggsave(
  boxplot_nutri,
  filename = "../images/boxplot_lab.png",
  width = 6,
  height = 5,
  dpi = 600
)
```

## Unir os gráficos

```{r fig.width= 6, fig.height=4}
library(patchwork)

boxplot_sonda <- (boxplot_wt + boxplot_od) /
  (boxplot_ph + boxplot_ce) +

  plot_layout(guides = "collect") &
  theme(legend.position = 'bottom')

boxplot_sonda

boxplot_sonda %>% ggsave(
  filename = "../images/boxplot_sonda.png",
  width = 6,
  height = 4,
  dpi = 600
)
```

```{r, fig.width=8, fig.height=10}
layout <- c(area(
  t = 1,
  l = 1,
  b = 2,
  r = 4
),
area(
  t = 3,
  l = 1,
  b = 4,
  r = 4
),
area(
  t = 5,
  l = 1,
  b = 11,
  r = 4
))

boxplot_geral <- (boxplot_wt + boxplot_od) /
  (boxplot_ph + boxplot_ce) /
  boxplot_nutri +
  plot_layout(guides = "collect", design = layout) &
  theme(legend.position = 'bottom')

boxplot_geral

boxplot_geral %>% ggsave(
  filename = "../images/boxplot_geral.png",
  width = 8,
  height = 10,
  dpi = 600
)

```